Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data

### 2021-08-02 Manuscript/SS_GAM.tex
Run pipeline

This is to leverage the fact that the probability of selecting the bases
of a smooth function should be similar, while allowing different penalty
on the null space and penalty space of the spline function. The hyper
prior of \(\gamma_{j}^{0}\) decides the sparsity of the model at the
function selection level, while that of \(\gamma_{j}^\text{pen}\)
decides the smoothness of the spline function at basis function level.
Meanwhile, we specify that \(\gamma_{j}^0\) and
\(\gamma_{j}^\text{pen}\) are independently distributed for analytic
simplicity. We further specify the parameter \(\theta_j\) follows a beta
distribution with given shape parameters \(a\) and \(b\), \[
\theta_j \sim Beta(a, b).
\] The beta distribution is a conjugate prior for the binomial
distribution and hence provides some computation convenience. For
simplicity, we focus on a special case of beta distribution, uniform
(0,1). When the variable have large effects in any of the bases, the
parameter \(\theta_j\) will be estimated large, which in turn encourages
the model to include the rest of bases, achieving the local adaption
among spline bases. Hereafter, we refer Bayesian hierarchical
generalized additive models with the spike-and-slab lasso prior as the
BHAM-SSL. \textbackslash{}

\begin{tikzpicture} [
staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
]

@@

With the re-parameterization step of the basis function matrix \(X\), it
is possible to generalized the SSL prior to other priors, for example
normal priors for ridge-type regularization and mixture normal prior for
spike-and-slab regularization. These priors would work better in low and
medium dimensional settings where the sparse assumption is not
necessary. Here we focus on the mixture normal prior as a demonstration
of applying continous spike-and-slab prior in additive models.
\footnote{TODO: enter a sentence introducing EMVS and spike-and-slab
  mixture normal prior}.

@@

@@

To evaluate the performance of the models, mean squared error, mean
absolute error for gaussian model, AUC and misclassifcation rate are
calculated for binomial model, and deviance are calculated for Poisson
model, averaging across 50 simulations. The results are tabulated in
Table 1, 2, 3.

\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-8-1.pdf}

@@

>>>>

This is to leverage the fact that the probability of selecting the bases
of a smooth function should be similar, while allowing different penalty
on the null space and penalty space of the spline function. The hyper
prior of \(\gamma_{j}^{0}\) decides the sparsity of the model at the
function selection level, while that of \(\gamma_{j}^\text{pen}\)
decides the smoothness of the spline function at basis function level.
Meanwhile, we specify that \(\gamma_{j}^0\) and
\(\gamma_{j}^\text{pen}\) are independently distributed for analytic
simplicity. We further specify the parameter \(\theta_j\) follows a beta
distribution with given shape parameters \(a\) and \(b\), \[
\theta_j \sim Beta(a, b).
\] The beta distribution is a conjugate prior for the binomial
distribution and hence provides some computation convenience. For
simplicity, we focus on a special case of beta distribution, uniform
(0,1). When the variable have large effects in any of the bases, the
parameter \(\theta_j\) will be estimated large, which in turn encourages
the model to include the rest of bases, achieving the local adaption
among spline bases. Hereafter, we refer Bayesian hierarchical
generalized additive models with the spike-and-slab lasso prior as the
BHAM-SSL, and visually presented in Figure \ref{fig:SSprior}.
\textbackslash{}

\begin{figure}
\centering
\begin{tikzpicture} [
staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
]

@@

With the re-parameterization step of the basis function matrix \(X\), it
is possible to generalized the SSL prior to other priors, for example
normal priors for ridge-type regularization and mixture normal prior for
spike-and-slab regularization. These priors would work better in low and
medium dimensional settings where the sparse assumption is not
necessary. Here we focus on the mixture normal prior as a demonstration
of applying continuous spike-and-slab prior in additive models.
\footnote{TODO: enter a sentence introducing EMVS and spike-and-slab
  mixture normal prior}.

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrllllll}
  \hline
 & p & bglm\_spline\_de & blasso\_spline & cosso & acosso & mgcv & SB\_GAM \\ 
  \hline
1 & 4.00 & 0.90 (0.01) & 0.90 (0.01) & 0.73 (0.09) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
  2 & 10.00 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.16) & 0.76 (0.04) & 0.90 (0.01) & 0.82 (0.04) \\ 
  3 & 50.00 & 0.88 (0.01) & 0.81 (0.05) & 0.43 (0.19) & 0.54 (0.19) & 0.86 (0.02) & 0.81 (0.07) \\ 
  4 & 200.00 & 0.73 (0.10) & 0.60 (0.08) & 0.39 (0.17) & 0.45 (0.19) & NaN (NA) & 0.81 (0.05) \\ 
   \hline
\end{tabular}
\caption{The average and standard deviation of the out-of-sample area under the curve measure over for binomial outcome over 50 iterations} 
\label{tab:gaus}
\end{table}

@@

To evaluate the predictive performance of the models, \(R^2\) for
Gaussian model, area under the curve (AUC) are calculated for binomial
model, averaging across 50 simulations. The results are tabulated in
Table \ref{tab:gaus}, \ref{tab:bin_auc}.\footnote{is it better to
  provide a box plot or a table.}

@@

\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-11-1.pdf}
